Learning to Synthesize Rules for Rule-based Intrusion Detectors

### 2020-03-06 paper/paper.tex
pass

\Fix{
denial-of-service
attack to a server by matching specific conditions about the network
traffic~\cite{understanding-dos}. Relevant properties about the
traffic of interest appear in bold in this rule (see
Section~\ref{sec:suri-metas-coverage}). Deployments of rule-based
\nids\ are restricted to a fixed set of rules defined by the network
system administrator.}\Mar{check...}

@@

Figure~\ref{fig:pingscan-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}.  A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The \emph{action} appears as the
first word in the rule description (an \CodeIn{alert}, in this
case). An action denotes the task that needs to be executed if the
rule pattern is satisfied.  In this example, an alert message will be
sent to system administrators if the rule pattern is
satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
  differs from an IDS in the action taken when an intrusion is
  observed. IPS systems block malicous traffic whereas IDS system only
  alert them. However, IDS and IPS do not differ in the approaches
  used to detect the intrusion~\cite{ids-ips}.\Mar{check}}. The
\emph{header} comes after the action in the rule description. It
restricts the information flow covered by the rule. The header of this
rule is \CodeIn{icmp any any -> any any}. It instructs Suricata to
inspect \CodeIn{icmp} traffic flowing from any machine and port to any
machine and port.\MyComment{ The variables \CodeIn{\$HOME\_NET} and
  \CodeIn{\$EXTERNAL\_NET} are configurable.} Finally, \emph{rule
  options} come after the header. It is a semi-colon-separated
sequence of key-value pairs. Rule options serve to document the
analyzed traffic (\eg{}, ``msg'', ``classtype'', and ``sid'') and to
characterize the attack pattern (\eg, ``flags'', ``threshold'', and
``content''). \tname{} produces options of the second kind, which are
\emph{relevant} to capture the attack. For example, the option
\CodeIn{content:<arg>} checks if the argument is present in the
payload of a message.

@@

The content strategy focuses on extracting content options from the
message payload. By inspecting existing rules, we found that these
options refer to sequences of characters separated by typical natural
language delimiters such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. To find the values associated with these
options, \tname{} splits the payload in tokens using these delimiters.
Considering the example from Section~\ref{sec:content-example},
\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
  Se sim, como isto cai para 5 (pela Tabela 3)}, considered content
option candidates.

@@

\subsubsection{Weight Optimization}
\label{sec:weight-optimization}
We optimized the weights of these heuristic functions as follows. We
ran \tname{} using different combinations of weights on a sample of
attacks and chose the combination that offered the best overall
performance, which we computed with the sum of the rankings of the
golden rule on each attack. The weight combination that led to the
smaller sum of positions (of the golden rule) in the rakings was
considered the best choice. We considered \emph{all} combinations of
the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
four different functions and five possible weight values, we needed to
run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
end of the process, the weight combination $\langle{}x,y,z,w\rangle$
was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
  pesos que a gente usar para cada funcao.} Although empirical results
showed that this strategy results in good performance, we remain to
evaluate other optimization strategies, such as evolutionary
alternatives as AVM~\cite{10.1109/32.57624}, to find better weights.

>>>>

\Fix{
denial-of-service
attack to a server by matching specific conditions about the network
traffic~\cite{understanding-dos}. Relevant properties about the
traffic of interest appear in bold in this rule (see
Section~\ref{sec:suri-metas-coverage}). Deployments of rule-based
\nids\ are restricted to a fixed set of rules defined by the network
system administrator.}

@@

Figure~\ref{fig:pingscan-example} shows an example rule of
Suricata~\cite{suricata}, a popular open-source \nids{} maintained by
the Open Information Security Foundation (OISF)~\cite{oisf}.  A
Suricata rule is divided in three parts---action, header, and rule
options~\cite{suri-rule-format}. The \emph{action} appears as the
first word in the rule description (an \CodeIn{alert}, in this
case). An action denotes the task that needs to be executed if the
rule pattern is satisfied.  In this example, an alert message will be
sent to system administrators if the rule pattern is
satisfied\footnote{Technically, an IPS (Intrusion Prevention System)
  differs from an IDS in the action taken when an intrusion is
  observed. IPS systems block malicous traffic whereas IDS system only
  alert them. However, IDS and IPS do not differ in the approaches
  used to detect the intrusion~\cite{ids-ips}.}. The
\emph{header} comes after the action in the rule description. It
restricts the information flow covered by the rule. The header of this
rule is \CodeIn{icmp any any -> any any}. It instructs Suricata to
inspect \CodeIn{icmp} traffic flowing from any machine and port to any
machine and port.\MyComment{ The variables \CodeIn{\$HOME\_NET} and
  \CodeIn{\$EXTERNAL\_NET} are configurable.} Finally, \emph{rule
  options} come after the header. It is a semi-colon-separated
sequence of key-value pairs. Rule options serve to document the
analyzed traffic (\eg{}, ``msg'', ``classtype'', and ``sid'') and to
characterize the attack pattern (\eg, ``flags'', ``threshold'', and
``content''). \tname{} produces options of the second kind, which are
\emph{relevant} to capture the attack. For example, the option
\CodeIn{content:<arg>} checks if the argument is present in the
payload of a message.

@@

The content strategy focuses on extracting content options from the
message payload. By inspecting existing rules, we found that these
options refer to sequences of characters separated by typical natural
language delimiters such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. To find the values associated with these
options, \tname{} splits the payload in tokens using these delimiters.
Considering the example from Section~\ref{sec:content-example},
\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
  Se sim, como isto cai para 5 (pela Tabela 3)} \Luc{6.1.1 eh o ataque 
JBoss. Na tabela constam 23 tokens.}, considered content
option candidates.

@@

\subsubsection{Weight Optimization}
\label{sec:weight-optimization}
We optimized the weights of these heuristic functions as follows. We
ran \tname{} using different combinations of weights on a sample of
attacks and chose the combination that offered the best overall
performance, which we computed with the sum of the rankings of the
golden rule on each attack. The weight combination that led to the
smaller sum of positions (of the golden rule) in the rakings was
considered the best choice. We considered \emph{all} combinations of
the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
four different functions and five possible weight values, we needed to
run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
end of the process, the weight combination $\langle{}x,y,z,w\rangle$
was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
  pesos que a gente usar para cada funcao.} \Luc{f1: 1, f2: 0.25, f3: 0.25, f4: 1} Although empirical results
showed that this strategy results in good performance, we remain to
evaluate other optimization strategies, such as evolutionary
alternatives as AVM~\cite{10.1109/32.57624}, to find better weights.
